---
title: "mvord: Supplementary examples"
author: "Rainer Hirk, Kurt Hornik and Laura Vana"
date: "`r Sys.Date()`"
output:
  html_vignette:
    toc: TRUE
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{mvord: Supplementary examples}
  %\VignetteDepends{mvord}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load package **mvord**
```{r}
library(mvord)
```

Load data set
```{r}
data("data_cr")
str(data_cr, vec.len = 3)
head(data_cr, n = 3)
```

## Example 1 - A simple example with 2 raters
A simple example with 2 raters (rater1 and rater2). The data set data_cr has a wide format and hence, we apply the multiple measurement object MMO2 on the left-hand side of the formula object. The covariates LR, LEV, PR, RSIZE, BETA are passed on the right-hand side of the formula. The thresholds are set equal by the argument thresholds.constraints.
```{r eval = FALSE}
res_cor_probit_2raters <- mvord(formula = MMO2(rater1, rater2) ~ 0 + LR + LEV + PR + RSIZE + BETA,
                               threshold.constraints = c(1, 1),
                               data = data_cr)
```

```{r echo = FALSE, results = 'hide', eval = TRUE}
cache <- TRUE
FILE <- "res_cor_probit_2raters.rda"
if (cache & file.exists(FILE)) {
  load(FILE)
} else {
  if (cache) {
res_cor_probit_2raters <- mvord(formula = MMO2(rater1, rater2) ~ 0 + LR + LEV + PR + RSIZE + BETA,
                               threshold.constraints = c(1, 1),
                               data = data_cr)
res_cor_probit_2raters <- mvord:::reduce_size.mvord(res_cor_probit_2raters)
  save(res_cor_probit_2raters, file  = FILE)
  } else {
      if(file.exists(FILE)) file.remove(FILE)
  }
}
```
The results are displayed by:
```{r}
summary(res_cor_probit_2raters)
```

The coefficients can be extracted by:
```{r}
coef(res_cor_probit_2raters)
```


## Example 2 - A simple example with 2 raters
A simple example with 3 raters (rater1, rater2 and rater3). The regression coefficients are set equal by the argument coef.constraints.
```{r eval = FALSE}
res_cor_probit_3raters <- mvord(formula = MMO2(rater1, rater2, rater 3) ~ 0 + LR + LEV + PR + RSIZE + BETA,
                               coef.constraints = c(1, 1, 1),
                               data = data_cr,
                               link = mvlogit())
```

```{r echo = FALSE, results = 'hide', eval = TRUE}
cache <- TRUE
FILE <- "res_cor_probit_3raters.rda"
if (cache & file.exists(FILE)) {
  load(FILE)
} else {
  if (cache) {
res_cor_probit_3raters <- mvord(formula = MMO2(rater1, rater2, rater3) ~ 0 + LR + LEV + PR + RSIZE + BETA,
                               coef.constraints = c(1, 1, 1),
                               data = data_cr,
                               link = mvlogit())
res_cor_probit_3raters <- mvord:::reduce_size.mvord(res_cor_probit_3raters)
  save(res_cor_probit_3raters, file  = FILE)
  } else {
      if(file.exists(FILE)) file.remove(FILE)
  }
}
```
The results are displayed by:
```{r}
summary(res_cor_probit_3raters)
```
The threshold parameters are presented by:
```{r}
thresholds(res_cor_probit_3raters)
```

## Example 3a - A model of firm ratings assigned by multiple raters and IG-SG indicator
This example presents a multivariate ordinal regression model with probit link and a general correlation error structure (cor_general(~ 1)).
```{r eval = FALSE}
res_cor_probit_simple <- mvord(formula = MMO2(rater1, rater2, rater3,
  rater4) ~ 0 + LR + LEV + PR + RSIZE + BETA, data = data_cr)
```

```{r echo = FALSE, results = 'hide', eval = TRUE}
cache <- TRUE
FILE <- "res_cor_probit_simple.rda"
if (cache & file.exists(FILE)) {
  load(FILE)
} else {
  if (cache) {
res_cor_probit_simple <- mvord(
    formula =
    MMO2(rater1, rater2, rater3, rater4) ~ 0 + LR + LEV + PR + RSIZE + BETA, data = data_cr)
res_cor_probit_simple <- mvord:::reduce_size.mvord(res_cor_probit_simple)
  save(res_cor_probit_simple, file  = FILE)
  } else {
      if(file.exists(FILE)) file.remove(FILE)
  }
}
```

The results are displayed by:
```{r}
summary(res_cor_probit_simple)
```
The threshold coefficients can be extracted by:
```{r}
thresholds(res_cor_probit_simple)
```
The regression coefficients are obtained by:
```{r}
coef(res_cor_probit_simple)
```

The error structure for firm with firm_id = 11 is displayed by:
```{r}
error_structure(res_cor_probit_simple)[[11]]
```


## Example 3b - A more elaborate model of ratings assigned by multiple raters
In this example, we extend the setting of the above example by imposing constraints on the
regression as well as on the threshold parameters and changing the link function to the
multivariate logit link.
```{r eval = FALSE}
res_cor_logit <- mvord(formula = MMO2(rater1, rater2, rater3, rater4) ~
    0 + LR + LEV + PR + RSIZE + BETA, data = data_cr, link = mvlogit(),
    coef.constraints = cbind(LR = c(1, 1, 1, 1), LEV = c(1, 2, 3, 4),
      PR = c(1, 1, 1, 1), RSIZE = c(1, 1, 1, 2), BETA = c(1, 1, 2, 3)),
    threshold.constraints = c(1, 1, 2, 3))
```

```{r echo = FALSE, results = 'hide', eval = TRUE}
FILE <- "res_cor_logit.rda"
if (cache & file.exists(FILE)) {
  load(FILE)
} else {
  if (cache) {
res_cor_logit <- mvord(formula = MMO2(rater1, rater2, rater3, rater4) ~
    0 + LR + LEV + PR + RSIZE + BETA, data = data_cr, link = mvlogit(),
  coef.constraints = cbind(
    c(1,1,1,1),
    c(1,2,3,4),
    c(1,1,1,1),
    c(1,1,1,2),
    c(1,1,2,3)),
    threshold.constraints = c(1, 1, 2, 3))
#res_cor_logit <- mvord:::reduce_size2.mvord(res_cor_logit)
  save(res_cor_logit, file  = FILE)
  } else {
      if(file.exists(FILE)) file.remove(FILE)
  }

}
```
The results are displayed by:
```{r}
summary(res_cor_logit)
```
The constraints are displayed by:
```{r}
constraints(res_cor_logit)
```
Note that the composite likelihood information criteria can be used for model comparison. For
objects of class ‘mvord’ CLAIC and CLBIC are computed by AIC() and BIC(), respectively.
The model fits of examples one and two are compared by means of BIC and AIC. We observe
that the model of Example 3b has a lower BIC and AIC indicating a better model fit:
```{r}
BIC(res_cor_probit_simple)
BIC(res_cor_logit)
```

```{r}
AIC(res_cor_probit_simple)
AIC(res_cor_logit)
```

The value of the pairwise log-likelihood of the two models can be extracted by:
```{r}
logLik(res_cor_probit_simple)
logLik(res_cor_logit)
```

Marginal predictions are obtained by:
```{r}
mp <- marginal_predict(res_cor_logit, type = "class")
```
```{r}
table(res_cor_logit$rho$y[,1], mp[,1])
table(res_cor_logit$rho$y[,2], mp[,2])
table(res_cor_logit$rho$y[,3], mp[,3])
table(res_cor_logit$rho$y[,4], mp[,4])
```

Joint predictions are obtained by:
```{r eval = FALSE}
jp <- joint_probabilities(res_cor_logit, type = "class")
```

```{r echo = FALSE, results = 'hide', eval = TRUE}
FILE <- "jp.rda"
if (cache & file.exists(FILE)) {
  load(FILE)
} else {
  if (cache) {
jp <- predict(res_cor_logit, type = "class")
  save(jp, file  = FILE)
  } else {
      if(file.exists(FILE)) file.remove(FILE)
  }
}
```

```{r}
table(res_cor_logit$rho$y[,1], jp[,1])
table(res_cor_logit$rho$y[,2], jp[,2])
table(res_cor_logit$rho$y[,3], jp[,3])
table(res_cor_logit$rho$y[,4], jp[,4])
```

## Example 4 - A longitudinal model
In this example, we present a longitudinal multivariate ordinal probit regression model
with a covariate dependent AR(1) error structure using the data set data_cr_panel:
```{r}
data(data_cr_panel)
str(data_cr_panel, vec.len = 3)
head(data_cr_panel, n = 3)
```


```{r eval=F}
res_AR1_probit <- mvord(formula = MMO(rating, firm_id, year) ~ LR + LEV +
  PR + RSIZE + BETA, error.structure = cor_ar1(~ BSEC), link = mvprobit(),
  data = data_cr_panel, coef.constraints = c(rep(1, 4), rep(2, 4)),
  threshold.constraints = rep(1, 8), threshold.values = rep(list(c(0, NA,
    NA, NA)),8), mvord.control(solver = "BFGS"))
```


```{r echo = FALSE, results = 'hide', eval = TRUE}
FILE <- "res_AR1_probit.rda"
if (cache & file.exists(FILE)) {
  load(FILE)
} else {
  if (cache) {
res_AR1_probit <- mvord(
  formula = MMO(rating, firm_id, year) ~ LR + LEV + PR + RSIZE +  BETA,
  data = data_cr_panel,
  error.structure = cor_ar1(~ BSEC),
  coef.constraints = c(rep(1,4), rep(2,4)),
  threshold.constraints = c(rep(1,8)),
  threshold.values = rep(list(c(0,NA,NA,NA)),8),
  link = mvprobit(),
  control = mvord.control(solver = "BFGS",  solver.optimx.control = list(trace = TRUE)))
   res_AR1_probit <- mvord:::reduce_size.mvord(res_AR1_probit)
  save(res_AR1_probit, file  = FILE)
  } else {
      if(file.exists(FILE)) file.remove(FILE)
  }
}
```
The results of the model can be presented by
```{r}
summary(res_AR1_probit, short = TRUE, call = FALSE)
```
The default error structure output for AR(1) models:
```{r}
error_structure(res_AR1_probit)
```
Correlation parameters ρ i for each firm are obtained by choosing type =
"corr"
```{r}
head(error_structure(res_AR1_probit, type = "corr"), n = 3)
```
Correlation matrices for each specific firm are obtained by choosing type =
"sigmas"
```{r}
head(error_structure(res_AR1_probit, type = "sigmas"), n = 1)
```
